---
title: "Reproducible Research Course  Peer Assessment 1"
author: "Edith Reyes Loya"
date: "18 de abril de 2015"
output: html_document
---

####*1. Loading and preprocessing the data*
```{r, echo=TRUE}
library(lubridate)
library(dplyr)
#Load the data (i.e. read.csv())
data<-read.csv("activity.csv")
#Process/transform the data (if necessary) into a format suitable for your analysis
data$date<-ymd(data$date)
str(data)
```


####*2. What is mean total number of steps taken per day?*
```{r, echo=TRUE}
suma<-tapply(data$steps, INDEX=as.factor(data$date), FUN=sum)
suma<-data.frame(suma)
head(suma)
#Calculate and report the mean and median total number of steps taken per day
media<-mean(suma$suma,na.rm=TRUE)
mediana<-median(suma$suma,na.rm=TRUE)
#Make a histogram of the total number of steps taken each day
hist(suma$suma,breaks=20,col="red",xlab="Intervals",main="Histogram of sum of steps per day")
abline(v=media,col="blue")
abline(v=mediana,col="purple")
text(x=4000,y=8,labels=c(paste("Mean = ",as.character(round(media,3),digits=3)),sep=" "),col="blue")
text(x=4000,y=7,labels=c(paste("Median = ",as.character(round(mediana,3),digits=3)),sep=" "),col="purple")
```

####*3.What is the average daily activity pattern?*
```{r, echo=TRUE}
medias<-tapply(data$steps, INDEX=data$interval, FUN=mean, na.rm=TRUE)
medias<-data.frame(medias)
#Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
maximo<-max(medias$medias)
k<-which.max(medias$medias)
cual<-row.names(medias)[k]
#Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-ax
plot(x=row.names(medias),y=medias$medias,type="l",xlab="intervals",ylab="means of every interval",col="blue",main="Means of every 5-minutes intervals")
abline(v=cual,col="red")
text(x=300,y=100,labels=c(paste("Maximum = ",as.character(round(maximo,3),digits=3)),sep=" "),col="red")
text(x=300,y=70,labels=c(paste("Interval = ",cual),sep=" "),col="red")
```

####*4.Imputing missing values*

For this point I review every mean data asking for NA values, and replace it with the mean that I previously calculated.
```{r, echo=TRUE}
k<-nrow(suma)
# Reviewing and reporting dates with missing values
for(i in 1:k) ifelse(is.na(suma$suma[i]),print(names(suma$suma[i])),NA)
nuevo<-data.frame(suma$suma)
colnames(nuevo)<-c("suma")
# Replacing the dates with missing values with media previously calculated
for(i in 1:k) nuevo$suma[i]=ifelse(is.na(suma$suma[i]),media,suma$suma[i])
nmedia<-mean(nuevo$suma)      # without na.rm
nmediana<-median(nuevo$suma)  # without na.rm
hist(nuevo$suma,breaks=20,col="red",xlab="Intervals",main="Histogram replacing NA's with mean")
abline(v=nmedia,col="blue")
abline(v=nmediana,col="purple")
text(x=4000,y=8,labels=c(paste("Mean = ",as.character(round(nmedia,3),digits=3)),sep=" "),col="blue")
text(x=4000,y=7,labels=c(paste("Median = ",as.character(round(nmediana,3),digits=3)),sep=" "),col="purple")
```

Do these values differ from the estimates from the first part of the assignment?
Yes, now the median is equal to the mean

What is the impact of imputing missing data on the estimates of the total daily number of steps?
Now the median is equal to the mean



####*5.Are there differences in activity patterns between weekdays and weekends?*
```{r, echo=TRUE}
#Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
weekday<-mutate(data,week=ifelse(weekdays(date)=="Sunday" | weekdays(date)=="Saturday","Weekend","Weekday"),interval,steps)
medias<-tapply(weekday$steps, INDEX=list(weekday$week,weekday$interval), FUN=mean, na.rm=TRUE)
columnas<-colnames(medias)
#Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
head(row.names(medias))
par(mfrow=c(2,1),mar=c(3,5,3,5))
plot(x=colnames(medias),y=medias[1,],type="l",xlab="intervals",ylab="means of every interval",col="blue",main=c("Means of every 5-minutes intervals in ",row.names(medias)[1]))
plot(x=colnames(medias),y=medias[2,],type="l",xlab="intervals",ylab="means of every interval",col="red",main=c("Means of every 5-minutes intervals in ",row.names(medias)[2]))
```