---
title: "Reproducible Research Course  Peer Assement 1"
author: "Edith Reyes Loya"
date: "18 de abril de 2015"
output: html_document
---

####*1. Loading and preprocessing the data*
```{r, echo=TRUE}

library(lubridate)
library(dplyr)
data<-read.csv("activity.csv")
data$date<-ymd(data$date)
str(data)
```


####*2. What is mean total number of steps taken per day?*
```{r, echo=TRUE}
suma<-tapply(data$steps, INDEX=as.factor(data$date), FUN=sum)
suma<-data.frame(suma)
head(suma)
media<-mean(suma$suma,na.rm=TRUE)
mediana<-median(suma$suma,na.rm=TRUE)
hist(suma$suma,breaks=20,col="red",xlab="Intervals",main="Histogram of sum of steps per day")
abline(v=media,col="blue")
abline(v=mediana,col="purple")
text(x=4000,y=8,labels=c(paste("Mean = ",as.character(round(media,3),digits=3)),sep=" "),col="blue")
text(x=4000,y=7,labels=c(paste("Median = ",as.character(round(mediana,3),digits=3)),sep=" "),col="purple")
```

####*3.What is the average daily activity pattern?*
```{r, echo=TRUE}
medias<-tapply(data$steps, INDEX=data$interval, FUN=mean, na.rm=TRUE)
medias<-data.frame(medias)
maximo<-max(medias$medias)
k<-which.max(medias$medias)
cual<-row.names(medias)[k]
plot(x=row.names(medias),y=medias$medias,type="l",xlab="intervals",ylab="means of every interval",col="blue",main="Means of every 5-minutes intervals")
abline(v=cual,col="red")
text(x=300,y=100,labels=c(paste("Maximum = ",as.character(round(maximo,3),digits=3)),sep=" "),col="red")
text(x=300,y=70,labels=c(paste("Interval = ",cual),sep=" "),col="red")
```

####*4.Imputing missing values*

For this point I review every mean data asking for NA values, and replace it with the mean that I previously calculated.
```{r, echo=TRUE}
k<-nrow(suma)
# Reviewing and reporting dates with missing values
for(i in 1:k) ifelse(is.na(suma$suma[i]),print(names(suma$suma[i])),NA)
nuevo<-data.frame(suma$suma)
colnames(nuevo)<-c("suma")
# Replacing the dates with missing values with media previously calculated
for(i in 1:k) nuevo$suma[i]=ifelse(is.na(suma$suma[i]),media,suma$suma[i])
nmedia<-mean(nuevo$suma)      # without na.rm
nmediana<-median(nuevo$suma)  # without na.rm
hist(nuevo$suma,breaks=20,col="red",xlab="Intervals",main="Histogram replacing NA's with mean")
abline(v=nmedia,col="blue")
abline(v=nmediana,col="purple")
text(x=4000,y=8,labels=c(paste("Mean = ",as.character(round(nmedia,3),digits=3)),sep=" "),col="blue")
text(x=4000,y=7,labels=c(paste("Median = ",as.character(round(nmediana,3),digits=3)),sep=" "),col="purple")
```

Do these values differ from the estimates from the first part of the assignment?
Yes, now the median is equal to the mean

What is the impact of imputing missing data on the estimates of the total daily number of steps?
Now the median is equal to the mean



####*5.Are there differences in activity patterns between weekdays and weekends?*
```{r, echo=TRUE}
weekday<-mutate(data,week=ifelse(weekdays(date)=="Sunday" | weekdays(date)=="Saturday","Weekend","Weekday"),interval,steps)
medias<-tapply(weekday$steps, INDEX=list(weekday$week,weekday$interval), FUN=mean, na.rm=TRUE)
columnas<-colnames(medias)
head(row.names(medias))
par(mfrow=c(2,1),mar=c(3,5,3,5))
plot(x=colnames(medias),y=medias[1,],type="l",xlab="intervals",ylab="means of every interval",col="blue",main=c("Means of every 5-minutes intervals in ",row.names(medias)[1]))
plot(x=colnames(medias),y=medias[2,],type="l",xlab="intervals",ylab="means of every interval",col="red",main=c("Means of every 5-minutes intervals in ",row.names(medias)[2]))
```